Discussion Draft

### 2020-03-17 doc/Discussion.rmd
better writing points

---
title: "Discussion"
date: ""
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document: default
  word_document:
    toc: yes
    toc_depth: '5'
bibliography: refs_Main_Report.bib
---

# Differences across models

Explain the model differences;

## 1.can it be applied to broad scale chromosome level landscape?

-genome wide recombination rates are made of the chromosome level patterns

This is important because general patterns have been observed for the recombination landscape at the chromosome level (also small range of crossovers per chromosome and broad scale pattern (Haenel et al ))

## 2. Does the model account for evolution?

given all the variance observed and fast evolution, models have to have some engine for evolution.

(some source of affecting fitness, or shifting the optima)

@@

## male telomeric, female centromeric Typical landscape
<review the typical landscape, one of the major broad scale patterns,
exceptions
consequences, (differences in variation, Sardell paper) >

   1. male telocentric and female uniform landscape / general placement of crossovers.  What is referecened to as typical landscape (Sardell) -- see references sardell kirkpatrick for list of species which show/display this typical phenotype

A difference in the general placement (general recombination landscape), with males having telomere bias and females having more uniform placement of crossovers along chromosomes.

-Exception to landscape (marisupials, birds?)

The Sardel guppies paper was able to deduce (dientangle) that this typical landscape pattern was due to // disentangle the specific components of centromere supression and telomere bias between the male and feamel landscapes. we are not able to do that in these cases (because of inability to estimate the chromosome identity across our sample of single bivalent
s. )

@@

1. SC length  2. connection to interference

@@

The pattern of interference across raw and normalized measures -- also confirms previous results of male having stronger interfere -- due to the longer loops (Petkov)

  2. Explain the connection between interference strength in physical space and (abstracted space)

>>>>

---
title: "Discussion Draft"
date: ""
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document: default
  word_document:
    toc: yes
    toc_depth: '5'
bibliography: refs_Main_Report.bib
---

```{r format.Notes, include=FALSE}

#table, wite by all of the columns and rows

#constrast on model types
#-evolution
#-models accounting for chromosome based traits

#HetC traits
# 1. typical landscape
# 2. female variance
# 3. chromatin organization

#faster male evolution traits
# 1. Reversed Direction of Heterochiasmy
# 2. correlated gwRR increase and interference strength

#go through each traits and point out which models support each of the traits

```

@@

## Broad Scale crossover placement patterns, male telomeric, female centromeric 'Typical landscape'

```{r typical.landscape.outline, include=FALSE}

#-long observed
#measured from a variety of approaches,
#larger variety of species

#-Exception to landscape (marisupials, birds?)

#which models make clear predctions for typical landscape?
#SACE
#drivers
#spindle differences

```

<review the typical landscape, one of the major broad scale patterns,
exceptions
consequences, (differences in variation, Sardell paper) >

The 'typical landscapes' refers to the sex differences in the board scale pattern or general placement of crossovers along the length of chromosomes. The primary pattern which has been observed across mny species is that crossovers are shifted to telomereic positions in males while a more unifrom landscape in observed in females [@sardell_sex_2020].

Our results show that the typical landscape is consered across all strains of house mice in this study (for all single chromosome daata and when controlled for crossover number), the broad scale position of crossovers was shifted closer to the telomere end (of the long arm).

<Detangling centromere suppression from telomere bias,  The Sardel guppies paper was able to deduce (dientangle) that this typical landscape pattern was due to // disentangle the specific components of centromere supression and telomere bias between the male and feamel landscapes. we are not able to do that in these cases (because of inability to estimate the chromosome identity across our sample of single bivalent
s. )>

@@

```{r chromatin.organization, include=FALSE}

#1. chromatin organization, female shorter loops, longer axis

#2. A consequence of chromatin organization (loop size) is difference in interference strength in (MB / cM)

#seen in humans, mice (signal comes through even with cell to cell variation in chromatin compaction / organization)
#

#exceptions to female pattern (fish birds), lack of data
# interference strength reversal (At and cattle) -- we predict that the loop size will also reverse, -- but there is a lack of cytological data for both sexes.

#lots of data on genetic maps for both sexes, but lack of cytological data to support

#our results, female longer axis, 
#review caveats -- why we believe the pattern is true

#refs
#BF klecner?
#negative relationship of axis and loop size (Lynn? Gruhn?)
#global re-orgnaiziton of DNA / chromosomes (3D, new Chip-seq pubs)
```

1. basics and expectations for chromatin organization

2. Well supported model for interference is that it is mediated / communicated through the 2D chromatin axis or SC (depending on the timing)

3. consequence for interference (general patterns) is that interference measured as (mb) will be stronger if DNA loops are stronger

The relationship // molecular organization of chromosome -- undergoes global re-orgzniation during meiosis. The basic take away is that eraly stage in meiosis) there is a central axis made up of chromatin and proteins while most of the DNA / chromatin are folded into loops that emenate off of the axis. This structure has been called 'lampbrush' by some (cite).

A pattern that has been long noticed --

Merging of two types of data

-total chromatin structure (axis length) (cytology)

-Interfere measured as cM or DNA base pairs Mb  (the spacing between markers from linkage map)

@@

**lack of difference in the raw measures of IFD, but significant difference in **

The pattern of interference across raw and normalized measures -- also confirms previous results of male having stronger interfere -- due to the longer loops (Petkov) **these results mirror those seen in** [@petkov2007]
